A general Nullstellensatz for generalized spaces

### 2021-11-06 paper-qcoh.tex
typo
to do

\begin{rem}From the internal point of view of~$\Set[\TT]$, we can construct the
classifying topos of~$\ul{\TT}/U_\TT$. Externally, this construction gives rise
to a bounded topos over~$\Set[\TT]$, hence to a Grothendieck topos. Using for
instance the technique described
in~\cite{blechschmidt-hutzler-oldenziel:composition}, one can show that this topos classifies the
theory of homomorphisms between~$\TT$-models. It can also be obtained as the
lax pullback~$(\Set[\TT] \Rightarrow_{\Set[\TT]} \Set[\TT])$.
There are two canonical geometric morphisms from this topos to~$\Set[\TT]$, the
morphism computing the domain and the morphism computing the codomain; the
morphism obtained by externalizing the internal construction is the former.
\end{rem}

@@

@@

Butz and Johnstone~\cite{butz-johnstone:first-order} list several nongeometric sequents holding in the
classifying topos of the theory~$\OO$ of objects, the theory with a single sort and
no function symbols, relation symbols or axioms, including:
\begin{align*}
  \forall x,y\?U_\OO\_ \neg\neg(x = y) &&
  \forall x_1,\ldots,x_n\?U_\OO\_ \neg \forall y\?U_\OO\_ \bigvee_{i=1}^n (y = x_i)
\end{align*}
Intuitively, the first expresses that~$U_\OO$ is close to a subsingleton, while
the second expresses that~$U_\OO$ is close to being infinite. Using the
Nullstellensatz, these formulas can be verified as follows.

We argue internally to~$\Set[\OO]$. Let~$x,y\?U_\OO$ and assume~$\neg(x = y)$.
Then~$\ul{\OO}/U_\OO$ proves~$(x = y \seq{[]} \bot)$. Hence this sequent holds
of any model of~$\ul{\OO}/U_\OO$, in particular of the
quotient~$U_\OO/(x \sim y)$. Hence~$\bot$ holds for this model, that is, we
have a contradiction.

@@

>>>>

\begin{rem}From the internal point of view of~$\Set[\TT]$, we can construct the
classifying topos of~$\ul{\TT}/U_\TT$. Externally, this construction gives rise
to a bounded topos over~$\Set[\TT]$, hence to a Grothendieck topos. Using for
instance the technique described
in~\cite{blechschmidt-hutzler-oldenziel:composition}, one can show that this topos classifies the
theory of homomorphisms between~$\TT$-models. This topos can also be obtained as the
lax pullback~$(\Set[\TT] \Rightarrow_{\Set[\TT]} \Set[\TT])$.
There are two canonical geometric morphisms from this topos to~$\Set[\TT]$, the
morphism computing the domain and the morphism computing the codomain; the
morphism obtained by externalizing the internal construction is the former.
\end{rem}

@@

% XXX explain how Nullstellensatz follows from the Horn version

@@

Butz and Johnstone~\cite{butz-johnstone:first-order} list several nongeometric sequents holding in the
classifying topos of the theory~$\OO$ of objects, the theory with a single sort and
no function symbols, relation symbols or axioms, including the following ones.
\begin{align*}
  \forall x,y\?U_\OO\_ \neg\neg(x = y) &&
  \forall x_1,\ldots,x_n\?U_\OO\_ \neg \forall y\?U_\OO\_ \bigvee_{i=1}^n (y = x_i)
\end{align*}
Intuitively, the first sequent expresses that~$U_\OO$ is close to a subsingleton, while
the second expresses that~$U_\OO$ is close to being infinite. Using the
Nullstellensatz, these formulas can be verified as follows.

We argue internally to~$\Set[\OO]$. Let~$x,y\?U_\OO$ and assume~$\neg(x = y)$.
By the Nullstellensatz, the theory~$\ul{\OO}/U_\OO$ proves~$(x = y \seq{[]} \bot)$. Hence this sequent holds
of any model of~$\ul{\OO}/U_\OO$, in particular of the
quotient~$U_\OO/(x \sim y)$. Hence~$\bot$ holds for this model, that is, we
have a contradiction.

@@

* XXX Cite C2.5.4 or similar
